---
title: "Data science for economists exams"
author: Jonathan
date: 04-22-2022
description: "Solving some fun problems"
image: "image_goes_here"
draft: true
format: 
  html:
    code-link: true
    code-overflow: wrap
---
# Purpose

The university of Oslo has produced a course called Data Science for Economists at the bachelor level. I want to complete the exams as a resource for my R screencasts, and because it looks fun!

## Exam 1

```{r}
library(tidyverse)
library(knitr)
theme_set(theme_light())

include_graphics(here::here("exams", "econ31_4170_2020h.pdf"))
```

### Question 1: Mobility under Covid-19

#### Subquestion 1a

```{r}
mobility <- read.csv(here::here("exams", "4170_vedlegg_2020_no_region_mobility_report.csv")) %>% 
  as_tibble()

mobility %>% skimr::skim()

mobility %>% count() -> n_obs
```

There are `r n_obs` observations.

```{r}
# transform into lubridate year month day format.
mobility <- mobility %>% 
  mutate(date = lubridate::ymd(date))

skimmed <- mobility %>% 
  skimr::skim()

skimmed$Date.min %>% as_tibble() %>% filter(!is.na(value)) -> date_min
skimmed$Date.max %>% as_tibble() %>% filter(!is.na(value)) -> date_max

```

The dates range from `r date_min` and `r date_max`.

#### Subquestion 1b

```{r}
mobility %>% count(sub_region_1, sort = T)
mobility %>% count(sub_region_2, sort = T)
```

UTF-8 can accommodate Norwegian characters like Ã.

#### Subquestion 1c

```{r}
oslo <- mobility %>% 
  filter(sub_region_1 == "Oslo")

oslo %>% 
  ggplot(aes(date, residential_percent_change_from_baseline)) +
  geom_point() +
  geom_smooth()
```

In Oslo, it looks like there was a big spike in March and a decline in movement over the summer, before an increase again towards Christmas.

There appear to be some day of the week effects also!

#### Subquestion 1d

```{r}
oslo <- oslo %>%
  mutate(
    weekday = lubridate::wday(date, label = T),
    weekend = case_when(
      weekday %in% c("Sat", "Sun") ~ 1,
      TRUE ~ 0
    ),
    weekend = factor(weekend)
  )


plot_oslo <- function(var) {
  oslo %>%
    ggplot(aes(date, residential_percent_change_from_baseline, colour = {{ var }})) +
    geom_point() +
    geom_smooth(se = F)
}

plot_oslo(weekday) +
  labs(colour = "Weekday")

plot_oslo(weekend) +
  labs(colour = "Is it a weekend?")
```

#### Subquestion 1e

Note that they store aggregate data in a strange way - if sub_region_2 is an empty string but sub_region_1 names a county then this is aggregate data at the county level.

```{r}
mobility %>%
  filter(sub_region_2 == "",
         sub_region_1 != "") %>% 
  mutate(
    weekday = lubridate::wday(date, label = T),
    weekend = case_when(
      weekday %in% c("Sat", "Sun") ~ "Weekend",
      TRUE ~ "Weekday"
    )
  ) %>% 
  ggplot(aes(date, residential_percent_change_from_baseline, colour = sub_region_1)) +
  geom_line() +
  # geom_smooth(se = F) +
  facet_wrap(~ weekend, nrow = 2)
```

Three weekday spikes are between April and June?

#### Subquestion 1f

Is this correct? Or should I be looking at squared differences?

```{r}
mobility %>%
  filter(
    sub_region_2 == "",
    sub_region_1 != "",
    between(date, lubridate::ymd("2020-03-12"), lubridate::ymd("2020-10-31"))
  ) %>% 
  group_by(sub_region_1) %>% 
  summarise(mean_response = mean(residential_percent_change_from_baseline, na.rm = T),
            median_response = median(residential_percent_change_from_baseline, na.rm = T)) %>% 
  ungroup() %>% 
  mutate(sub_region_1 = fct_reorder(sub_region_1, mean_response)) %>% 
  pivot_longer(-sub_region_1) %>% 
  mutate(value = round(value, 2)) %>% 
  ggplot(aes(value, sub_region_1, fill = name)) +
  geom_col() +
  geom_text(aes(label = value), hjust = 1.2) +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~ name) +
  theme(legend.position = "none") +
  labs(x = "Summary of residential percent change from baseline")
```

#### Subquestion 1g

```{r}
national <- mobility %>% 
  filter(sub_region_1 == "")

oslo_compare <- national %>% 
  select(residential_percent_change_from_baseline, date) %>% 
  rename(national = residential_percent_change_from_baseline) %>% 
  bind_cols(oslo %>% select(residential_percent_change_from_baseline) %>% 
              rename(oslo = residential_percent_change_from_baseline))

oslo_compare %>% 
  pivot_longer(-date) %>% 
  ggplot(aes(date, value, colour = name)) +
  geom_line()

oslo_compare %>% 
  mutate(week = lubridate::week(date)) %>% 
  group_by(week) %>% 
  mutate(across(c(national, oslo), mean)) %>% 
  ungroup() %>% 
  distinct(national, oslo, week) %>% 
  mutate(oslo_national = oslo / national) %>% 
  ggplot(aes(week, oslo_national)) +
  geom_line()

```

???

### Question 2

#### Subquestion 2a

```{r}
df <- tibble(x = 1:10,
             y = 1:10)

gamma = .5
theta = .3
phi = 2

df %>% 
  mutate(comp_1 = ((x^(1-gamma)/1-gamma)),
         comp_2 = phi*((y^(1-theta)/1-theta)),
         utility = comp_1 + comp_2) %>% 
  pivot_longer(-c(x, y)) %>% 
  ggplot(aes(x, value, colour = name)) +
  geom_point() +
  geom_line()
```


### Question 3

#### Subquestion 3a

Axis is not from zero! Sneaky!

```{r}
df <- tibble(chrome = 29.619,
             edge = 31.786,
             firefox = 26.876)

df %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(name = fct_reorder(name, value)) %>% 
  ggplot(aes(value, name, fill = name)) +
  geom_col(show.legend = F) +
  labs(x = "Speed score",
       y = NULL)

```

#### Subquestion 3b

Similar!

```{r}
df <- tibble(Latvia = 5.5,
             Australia = 5.4,
             Scotland = 5.4,
             Peru = 5.4,
             `South Africa` = 5.2,
             India = 5)

df %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(name = fct_reorder(name, value)) %>% 
  ggplot(aes(value, name, fill = name)) +
  geom_col(show.legend = F) +
  labs(x = "Average female height",
       y = NULL)

```


